---
title: "EFFECTS OF PARTICIPATORY MONITORING AND EVALUATION ON THE SUSTAINABILITY OF ECONOMIC STIMULUS PROGRAM IN THE KENYAN EDUCATION SECTOR: A CASE OF NAIROBI COUNTY, KENYA"
author: "Lucy"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  echo: false
  warning: false
fig-width: 7
fig-height: 5
---

```{r}
library(gt)
library(readxl)
library(tidyverse)
library(gtsummary)
library(showtext)
library(glue)
library(scales)
font_add_google(name = "Noto Serif", family = "Noto serif")
showtext_auto()
lucy <- read_excel("lucy-dataset.xlsx") |> 
  mutate(Gender = if_else(Gender == "male", "Male", "Female"))
```

## Social Demographics
```{r}
lucy |> select(Gender:Working) |> 
  mutate(Age = fct_relevel(Age, c("below 20", "20-29", "30-39", "40-49","50 years and above")),
         Education = fct_relevel(Education, c("Diploma","Bachelor's Degree", "Masters Degree", "PHD", "Others")),
         Working = fct_relevel(Working, c("-1 year", "2-5 years", "5-10 years", "10-15 years","15-20 years", "20+")))  |> 
  tbl_summary() |> as_gt()
```

## Gender
```{r}
Gender <- lucy |> count(Gender) |> 
  mutate(percent = round(n/sum(n)*100),
         Label = glue("{Gender}: {percent}%"))

pie(x=Gender$n, labels = Gender$Label, family="Noto serif")
```

## Age
```{r}
Age <- lucy |> count(Age) |> 
  mutate(percent = round(n/sum(n)*100),
         Label = glue("({Age}): {percent}%"))

pie(x=Age$n, labels = Age$Label, family="Noto serif")
```


## Education Level
```{r}
Education <- lucy |> count(Education) |> 
  mutate(percent = round(n/sum(n)*100),
         Label = glue("({Education}): {percent}%"))

pie(x=Education$n, labels = Education$Label, family="Noto serif")
```

## Working Experience
```{r}
Working <- lucy |> count(Working) |> 
  mutate(percent = round(n/sum(n)*100),
         Label = glue("({Working}): {percent}%"))

pie(x=Working$n, labels = Working$Label, family="Noto serif")
```

# Stakeholder Involvement
```{r}
lucy |> select(PME_awareness:PME_interactions) |> 
  tbl_summary()
```

## PME Awareness pie
```{r}
Aware <- lucy |> count(PME_awareness) |> 
  mutate(percent = round(n/sum(n)*100),
         Label = glue("({PME_awareness}): {percent}%"))

pie(x=Aware$n, labels = Aware$Label, family="Noto serif")
```

## PME Awareness barplot
```{r}
lucy |> 
  ggplot(aes(PME_awareness)) + geom_bar(stat = "count") +
  theme_classic(base_size = 20, base_family = "Noto serif") +
  labs(y = "Respondents Involved") +
  scale_y_continuous(breaks = seq(0, 55, 11))
```


## PME use pie
```{r}
use <- lucy |> count(`PME use`) |> 
  mutate(percent = round(n/sum(n)*100),
         Label = glue("({`PME use`}): {percent}%"))

pie(x=use$n, labels = use$Label, family="Noto serif")
```

## PME use barplot
```{r}
lucy |> 
  ggplot(aes(`PME use`)) + geom_bar(stat = "count") +
  theme_classic(base_size = 20, base_family = "Noto serif") +
  labs(y = "Respondents Involved") +
  scale_y_continuous(breaks = seq(0, 55, 11)) +
  scale_x_discrete(labels = label_wrap(10))
```


## PME interaction pie
```{r}
INTER <- lucy |> count(PME_interactions) |> 
  mutate(percent = round(n/sum(n)*100),
         Label = glue("({PME_interactions}): {percent}%"))

pie(x=INTER$n, labels = INTER$Label, family="Noto serif")
```

## PME use barplot
```{r}
lucy |> count(PME_interactions) |> 
  ggplot(aes(fct_reorder(PME_interactions, n), n)) + geom_bar(stat = "identity") +
  theme_classic(base_size = 20, base_family = "Noto serif") +
  labs(y = "Respondents Involved", x = "PME interactions") +
  scale_y_continuous(breaks = seq(0, 50, 10)) +
  scale_x_discrete(labels = label_wrap(10))
```

# Key indicators of Stakeholder Involvement
## Categorical wise
```{r}
lucy |> select(`Citizen control`:`Regular Debriefing`) |> 
  pivot_longer(cols = everything()) |> 
  mutate(value = str_to_sentence(value),
         value = if_else(value == "Srongly agree", "Strongly agree", value),
         value = fct_relevel(value, c("Strongly agree", "Agree", "Neutrol", "Disagree", "Strongly disagree"))) |> 
  tbl_summary(by = value,
              percent = "row")
```

## Interval wise
```{r}
lucy |> select(`Citizen control`:`Regular Debriefing`) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutrol" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  rowwise() |> 
  mutate(indicators = mean(c_across(1:8)))|> gt()

```

# indicators 2 
```{r}
lucy |> select(`Citizen control`:`Regular Debriefing`) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutrol" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  pivot_longer(cols = everything()) |> 
  summarise(Mean = mean(value),
            Sd = sd(value),
            .by = name)|> gt()

```

## Capacity building 1
### categorical wise
```{r}
lucy |> select(capacity_building_1:capacity_building_3) |> 
  mutate_all(~str_to_sentence(.)) |> 
  pivot_longer(cols = everything()) |> 
  mutate(value = str_squish(value),
         value = fct_relevel(value,
                             c("Strongly agree", "Agree",
                               "Neutrol", "Disagree", "Strongly disagree"))) |> 
  tbl_summary(by = value,
              percent = "row")
```

## Cpacity building interval wise
```{r}
lucy |> select(capacity_building_1:capacity_building_3) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutrol" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  rowwise() |> 
  mutate(capacity = mean(c_across(1:3))) |> gt()
```

## Capcity 2
```{r}
lucy |> select(capacity_building_1:capacity_building_3) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutrol" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  pivot_longer(cols = everything()) |> 
  summarise(Mean = mean(value),
            Sd = sd(value),
            .by = name)|> gt()
```

## Stakeholder ownership
### Categorical wise
```{r}
lucy |> select(Stakeholder_ownership_1:Stakeholder_ownership_4) |> 
  mutate_all(~str_to_sentence(.)) |> 
  pivot_longer(cols = everything()) |> 
  mutate(value = if_else(value == "Neutrol", "Neutral", value),
         value = fct_relevel(value,
                             c("Strongly agree", "Agree",
                               "Neutral", "Disagree", "Strongly disagree"))) |> 
  tbl_summary(by = value,
              percent = "row")
```

### Stakeholder ownership interval wise
```{r}
lucy |> select(Stakeholder_ownership_1:Stakeholder_ownership_4) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutral" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  rowwise() |> 
  mutate(capacity = mean(c_across(1:4)))|> gt()
```

### stakeholder 2
```{r}
lucy |> select(Stakeholder_ownership_1:Stakeholder_ownership_4) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutral" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  pivot_longer(cols = everything()) |> 
  summarise(Mean = mean(value),
            Sd = sd(value),
            .by = name)|> gt()
```

## Policy recommendation
## Categorical wise
```{r}
lucy |> select(`Policy recommendation_1`:`Policy recommendation_3`) |> 
  mutate_all(~str_to_sentence(.)) |> 
  pivot_longer(cols = everything()) |> 
  mutate(value = if_else(value == "Neutrol", "Neutral", value),
         value = fct_relevel(value,
                             c("Strongly agree", "Agree",
                               "Neutral", "Disagree", "Strongly disagree"))) |> 
  tbl_summary(by = value,
              percent = "row")
```

## Interval wise

```{r}
lucy |> select(`Policy recommendation_1`:`Policy recommendation_3`) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutral" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  rowwise() |> 
  mutate(capacity = mean(c_across(1:3)))
```

### recommendation 2
```{r}
lucy |> select(`Policy recommendation_1`:`Policy recommendation_3`) |> 
  mutate_all(~str_to_sentence(.)) |> 
  mutate_all(~case_when(.== "Strongly agree" ~ 5,
                        .== "Agree" ~ 4,
                        .== "Neutral" ~ 3,
                        .== "Disagree" ~ 2,
                        .== "Strongly disagree" ~ 1)) |> 
  mutate_all(~replace_na(.,1)) |> 
  pivot_longer(cols = everything()) |> 
  summarise(Mean = mean(value),
            Sd = sd(value),
            .by = name)|> gt()
```

