---
title: "Juvenile Anguillid Corrections"
author: "Lena"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  echo: false
  warning: false
---

```{r}
library(readxl)
library(tidyverse)
library(gtsummary)
library(gt)
library(readxl)
library(glue)
library(RColorBrewer)
library(showtext)
library(janitor)
font_add_google(name = "Noto Serif", family = "Noto serif")
showtext_auto()
Anguillids <- read_excel("Juvenile Anguillid.xlsx", 
    sheet = "Combined Juveniles") |> janitor::clean_names() |> 
  select(-c(effort))

Juvenile <- read_excel("Juvenile Anguillid.xlsx", 
    sheet = " Juvenile Only") |> 
  mutate(Season = if_else(str_detect(Season, "Wet"), "Wet", "Dry"),
         Total_Length = `TL (cm)` *10,
         `DG (cm)` = `DG (cm)` *10,
         Species = case_when(str_detect(Species, "benga") ~ "Anguilla bengalensis",
                             str_detect(Species, "mos") ~ "Anguilla mossambica",
                             .default = "Anguilla marmorata"),
         Month = str_to_sentence(Month)) |> 
  select(-c(Effort, Technique, `TL (cm)`)) |> 
  janitor::clean_names()
```
## All anguillids
```{r}
Anguillids |> 
  summarise(total_number = sum(n, na.rm = T),
            across(
              .cols = c(total_length_cm:weight_g),
              ~ mean(.,na.rm = T)),
            .by = c(family, species)) |> 
  gt(rowname_col = "species",
     groupname_col = "family") |> 
  tab_spanner(label = "Morphometric Measurements",
              columns = c(total_number:weight_g))
```




## Morphometrics
```{r}
Juvenile |> 
  select(where(is.numeric), species) |> 
  select(total_length, everything()) |> 
  tbl_summary(by = species,
              statistic = everything() ~ "{mean} ± {sd}") |> 
  add_p()
```

## Elvers Morhometrty
```{r}
Juvenile |> 
  select(where(is.numeric), species, local_name) |> 
  filter(local_name == "Elver") |> 
  select(total_length, everything(), -local_name) |> 
  tbl_summary(by = species,
              type = everything() ~ "continuous",
              statistic = all_continuous() ~ "{mean} ± {sd}") |> 
  add_p()
```

## Glass Eels Morhometrty
```{r}
Juvenile |> 
  select(where(is.numeric), species, local_name) |> 
  filter(local_name == "Glass eel") |> 
  select(total_length, everything(), -local_name) |> 
  tbl_summary(by = species,
              type = everything() ~ "continuous",
              statistic = all_continuous() ~ "{mean} ± {sd}") |> 
  add_p()
```

## Anguillids across seasons
```{r}
Juvenile |> 
  select(season, species) |> 
  tbl_cross(percent = "row") |> add_p()
```


## Elvers Across Seasons
```{r}
Juvenile |> 
  select(local_name, season, species) |> 
  filter(local_name == "Elver")|> 
  select(-local_name) |> 
  tbl_cross(percent = "row")
```


## Glass Eels Across Seasons
```{r}
Juvenile |> 
  select(local_name, season, species) |> 
  filter(local_name == "Glass eel")|> 
  select(-local_name) |> 
  tbl_cross(percent = "row") |> add_p()
```

# Phases of the moon
## Anguillids across moon phase
```{r}
Juvenile |> 
  select(moon_phase, species) |> 
  tbl_cross(percent = "row") |> add_p()
```


## Elvers Across moon phase
```{r}
Juvenile |> 
  select(local_name, moon_phase,species) |> 
  filter(local_name == "Elver")|> 
  select(-local_name) |> 
  tbl_cross(percent = "row")
```


## Glass Eels Across moon phase
```{r}
Juvenile |> 
  select(local_name, moon_phase,species) |> 
  filter(local_name == "Glass eel")|> 
  select(-local_name) |> 
  tbl_cross(percent = "row") |> add_p()
```

# Temporal Distribution
```{r}
Juvenile |> select(month, species) |> 
  mutate(month = fct_relevel(month, 
                             c("July", "August", "September", "October", "November"))) |> 
  tbl_cross(percent = "row") |> add_p()
  
```

## Tempral Distribution of elvers
```{r}
Juvenile |> select(month, local_name, species) |> 
  filter(local_name == "Elver") |> 
  select(-local_name) |> 
  tbl_cross(percent = "row")
```


## Tempral Distribution of glass eels
```{r}
Juvenile |> select(month, local_name, species) |> 
  filter(local_name == "Glass eel") |> 
  select(-local_name) |> 
  mutate(month = fct_relevel(month, 
                             c("July", "August", "September", "October", "November"))) |> 
  tbl_cross(percent = "row") |> add_p()
```

## Juveniles Plot
```{r}
Juvenile |> 
  count(species, local_name) |> 
  ggplot(aes(local_name, n, fill = species)) + geom_col(position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Juvenile Stage", y = "Juvenile Counts") +
  theme_bw(base_family = "Noto serif", base_size = 20) +
  theme(legend.position = c(0.8, 0.8)) 
```
## Biotopes
```{r}
Juvenile |> select(biotope, species) |> 
  tbl_cross(percent = "row") |> add_p()
```

## EELS BIOTOPES
```{r}
Juvenile |> select(biotope, local_name, species) |> 
  filter(local_name == "Glass eel") |> 
  select(-local_name) |> 
  tbl_cross(percent = "row") |> add_p()
```

## elvers BIOTOPES
```{r}
Juvenile |> select(biotope, local_name, species) |> 
  filter(local_name == "Elver") |> 
  select(-local_name) |> 
  tbl_cross(percent = "row") |> add_p()
```
## Catch Per Unit Effort
```{r}
Juvenile |> 
  mutate(date = convert_to_date(date, character_fun = lubridate::dmy)) |> 
  count(local_name,species) |> 
  pivot_wider(names_from = local_name, values_from = n) |> 
  mutate(Total = Elver + `Glass eel`) |> gt()
```


