---
title: "Insurance Models"
author: "Kuei"
format: pptx
editor: visual
editor_options: 
  chunk_output_type: console
fig-width: 9
fig-height: 5
---

```{r}
library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(showtext)
font_add_google(name = "Noto Serif", family = "Noto serif")
showtext_auto()
Data <- read_excel("Insurance.xlsx") %>% 
  mutate(Age = fct_relevel(Age, c("Below 30 years", "30 - 39 years",
                                  "40 - 49 years", "Over 50 years")),
         `Level of Education`  = fct_relevel(`Level of Education`, c("Secondary","Diploma", 
                                                                     "Bachelor's", "Postgraduate")),
         `Working Period` = fct_relevel(`Working Period`, c("Less than 1 year", "1 - 5 years",
                                                            "6 - 10 years", "over 10 years")),
         `Company Operation Period` = fct_relevel(`Company Operation Period`, c("Less than 5 years",
                                                                                "5 - 10 years",
                                                                                "Over 10 years")))
```

# Social Demographics

## Demographics summary

```{r}
Data %>% select(1:5) %>% 
  tbl_summary() %>% 
  modify_header(label = "**Social Demographics**",
                stat_0 = "**Frequency (%)**")
```

## Gender

```{r}
plot_demos <- function(var) {
  Data %>% 
  ggplot(aes({{var}})) + geom_bar(stat = "count") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(y = "Respondents")
}

plot_demos(Gender)
```

## Age

```{r}
plot_demos(Age) 
```

## Level of education

```{r}
plot_demos(`Level of Education`) 
```

## Occupation

```{r}
plot_demos(Occupation) 
```

## Working Period

```{r}
plot_demos(`Working Period`) 
```

## Company Operation Period

```{r}
plot_demos(`Company Operation Period`) 
```



## Overall Model

```{r}
library(performance)
model <- lm(`Financial Performance` ~ `Financial innovation` + `Longterm Investment` +
              `Equity Investment` + `Asset Base` + Gender + Age + `Working Period` + `Company Operation Period` + `Level of Education` + Occupation, data = Data)

summary(model)
tbl_regression(model)
check_model(model)
model_performance(model) %>% gt()
```

# Objective 1: The effect of financial innovations on financial performance

```{r}
model_1 <- lm(`Financial Performance` ~ `Financial innovation`, data = Data)

summary(model_1)
tbl_regression(model_1)
check_model(model_1)
model_performance(model_1) %>% gt()
```

# Objective 2: The effect of long- term investment on financial performance

```{r}
model_2 <- lm(`Financial Performance` ~ `Longterm Investment`, data = Data)

summary(model_2)
tbl_regression(model_2)
check_model(model_2)
model_performance(model_2) %>% gt()
```

#Objective 3: The relationship between equity investment and financial performance

```{r}
model_3 <- lm(`Financial Performance` ~  `Equity Investment`, data = Data)

summary(model_3)
tbl_regression(model_3)
check_model(model_3)
model_performance(model_3) %>% gt()
```
# Objective 4 :The moderating effect of asset base relationship between financial management determinants and performance

```{r}
model_4 <- lm(`Financial Performance` ~  `Financial innovation`*`Asset Base` +
                `Longterm Investment`*`Asset Base` +
                `Equity Investment`*`Asset Base`, data = Data)

summary(model_4)
tbl_regression(model_4)
check_model(model_4)
anova(model_4) %>% gt()
model_performance(model_4) %>% gt()
```
