---
title: "ASSESSMENT OF GENERALIZED ANXIETY DISORDER AND COMMUNITY-BASED INTERVENTIONS"
author: "Manishiwe"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  echo: false
  warning: false
---

```{r}
library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
library(showtext)
library(RColorBrewer)
library(glue)
library(webr)
font_add_google(name = "Noto Serif", family = "Noto serif")
showtext_auto()
Anxiety <- read_excel("Anxiety.xlsx") %>% 
  mutate(`Marital status` = fct_relevel(`Marital status`, c("Single", "Married", 
                                                            "Widowed", "Single mother")))

Anxiety_clean <- read_excel("Anxiety_clean.xlsx")
library(ggpubr)
```

# Social Demographics

## Age

```{r}
#| fig-width: 11
#| fig-height: 7
Anxiety %>% count(Age) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(Age, percent)) + geom_bar(stat = "identity") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(y = "Respondents") +
  scale_y_continuous(breaks = seq(0, 1, 0.05), labels = scales::percent_format())
```

## Gender

```{r}
#| fig-width: 11
#| fig-height: 7
Anxiety %>% count(Gender) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(Gender, percent)) + geom_bar(stat = "identity") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(y = "Respondents") +
  scale_y_continuous(labels = scales::percent_format())
```

## Marital Status

```{r}
#| fig-width: 11
#| fig-height: 7
Anxiety %>% count(`Marital status`) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(`Marital status`, percent)) + geom_bar(stat = "identity") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(y = "Respondents") +
  scale_y_continuous(breaks = seq(0,1,0.09),labels = scales::percent_format())
```

## Level of education

```{r}
#| fig-width: 11
#| fig-height: 7
Anxiety %>% count(`Education Level`) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(`Education Level`, percent)) + geom_bar(stat = "identity") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(y = "Respondents") +
  scale_y_continuous(labels = scales::percent_format())
```

## Occupation

```{r}
#| fig-width: 11
#| fig-height: 7
Anxiety %>% count(Occupation) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(percent, fct_reorder(Occupation, percent))) + geom_bar(stat = "identity") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(x = "Respondents", y = NULL) +
  scale_x_continuous(breaks = seq(0,1, 0.1), labels = scales::percent_format())
```

## Residence

```{r}
#| fig-width: 11
#| fig-height: 7
Anxiety %>% count(`Residence period`) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(`Residence period`,percent)) + geom_bar(stat = "identity") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(y = "Respondents") +
  scale_y_continuous(breaks = seq(0, 1, 0.1), labels = scales::percent_format())
```

## Nationality

```{r}
#| fig-width: 11
#| fig-height: 7
Anxiety %>% count(Nationality) %>% 
  mutate(percent = n/sum(n)) %>%
  ggplot(aes(percent, fct_reorder(Nationality, percent))) + geom_bar(stat = "identity") +
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(x = "Respondents", y = NULL) +
  scale_x_continuous(breaks = seq(0, 1, 0.10), labels = scales::percent_format())
```

## Social Demographics summary

```{r}
Anxiety %>% select(Age:Nationality) %>% 
  tbl_summary() %>% 
  modify_header(label = "**Social Demographic Characteristics**",
                stat_0 = "**Frequency (%)**")
  
```

# Anxiety

```{r}
GAD <- bind_cols(Anxiety %>% select(1:7),
          Anxiety %>% 
  select(GAD1:GAD7) %>% 
  mutate_all(~case_when(. == "Not at all" ~ 0,
                        . == "Several days" ~ 1,
                        . == "More than half the days" ~ 2,
                        . == "Nearly every day" ~ 3)) %>% 
  rowwise() %>% 
  mutate(GAD_SCORE = sum(c_across(GAD1:GAD7), na.rm = T)) %>% ungroup() %>% 
  mutate(GAD_LEVEL = case_when(between(GAD_SCORE, 0, 4) ~ "Minimal Anxiety",
                               between(GAD_SCORE, 5, 9) ~ "Mild Anxiety",
                               between(GAD_SCORE, 10, 14) ~ "Moderate Anxiety",
                               between(GAD_SCORE, 15, 21) ~ "Severe Anxiety")) %>% 
  select(GAD_SCORE, GAD_LEVEL))%>% 
  mutate(GAD_LEVEL = fct_relevel(GAD_LEVEL, c("Minimal Anxiety", "Mild Anxiety",
                                              "Moderate Anxiety", "Severe Anxiety"))) 
```

## Prevalence of General Anxiety Disorder

```{r}
GAD %>% select(GAD_LEVEL) %>% 
  tbl_summary() %>% 
  modify_header(label = "**Level General Anxiety Disorder**",
                stat_0 = "**Prevalance of General Anxiety Disorder**")
```

## Prevalence viz

```{r}
GAD %>% count(GAD_LEVEL) %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(GAD_LEVEL, percent)) + geom_col() + 
  theme_classic(base_size = 22, base_family = "Noto serif") +
  labs(x = "General Anxiety Disorder Levels", y = "Prevalence") +
  scale_y_continuous(breaks = seq(0, 1, 0.09), labels = scales::percent_format())
  
```

## Association between General disorder and social demographics

```{r}
GAD %>% select(-GAD_SCORE) %>% 
  tbl_summary(by = GAD_LEVEL) %>% 
  add_p()
```

## Comparisons between social demographics and anxiety scores

### Age

```{r}
GAD %>% select(Age, GAD_SCORE) %>% 
  tbl_summary(by = Age, 
              statistic = GAD_SCORE ~ "{mean} ({sd})")%>% 
  add_p()
```

### Gender

```{r}
GAD %>% select(Gender, GAD_SCORE) %>% 
  tbl_summary(by = Gender, 
              statistic = GAD_SCORE ~ "{mean} ({sd})")%>% 
  add_p()
```

### Marital status

```{r}
GAD %>% select(`Marital status`, GAD_SCORE) %>% 
  tbl_summary(by = `Marital status`, 
              statistic = GAD_SCORE ~ "{mean} ({sd})")%>% 
  add_p()
```

### Education Lecwl

```{r}
GAD %>% select(`Education Level`, GAD_SCORE) %>% 
  tbl_summary(by = `Education Level`, 
              statistic = GAD_SCORE ~ "{mean} ({sd})")%>% 
  add_p()
```

### Occupation

```{r}
GAD %>% select(Occupation, GAD_SCORE) %>% 
  tbl_summary(by = Occupation, 
              statistic = GAD_SCORE ~ "{mean} ({sd})")%>% 
  add_p()
```

### Residence Period

```{r}
GAD %>% select(`Residence period`, GAD_SCORE) %>% 
  tbl_summary(by = `Residence period`, 
              statistic = GAD_SCORE ~ "{mean} ({sd})")%>% 
  add_p()
```

### Nationality

```{r}
GAD %>% select(Nationality, GAD_SCORE) %>% 
  tbl_summary(by = Nationality, 
              statistic = GAD_SCORE ~ "{mean} ({sd})") %>% 
  add_p()


```

# Precipitating Factors

```{r}
Anxiety_clean %>% 
  select(Uncertainty:`no income`) %>% 
  mutate_all(~if_else(.== 1, "Not Causing Anxiety", "Causing Anxiety")) %>% 
  tbl_summary(missing = "no")
```

# Community based interventions

```{r}
Anxiety_clean %>% 
  select(`Individual counselling`:`job opportunities`) %>% 
  mutate_all(~if_else(between(.,0,1), "Not Satisfied", "Satisfied")) %>% 
  tbl_summary(missing = "no")
```

## Association between social demogrpahics and anxiety

## Age

```{r}
demo_tab <- function(demographic) {
 GAD %>% 
  select({{demographic}}, GAD_LEVEL) %>% 
  tbl_cross(percent = "column") 
}

demo_tab(Age)
```

## Gender

```{r}
demo_tab(Gender)
```

## Marital status

```{r}
demo_tab(`Marital status`)
```

## Education level

```{r}
demo_tab(`Education Level`)
```

## Occupation

```{r}
demo_tab(Occupation)
```

## Residence period

```{r}
demo_tab(`Residence period`)
```

## Nationality

```{r}
demo_tab(Nationality)
```

## Pie charts

### Age

```{r}
#| fig-width: 11
#| fig-height: 7
Age <- GAD %>% count(Age) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{Age}: {percent}%"))

pie(x = Age$percent, labels = Age$label, family = "Noto serif")
```

### Gender

```{r}
#| fig-width: 11
#| fig-height: 7
Gender <- GAD %>% count(Gender) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{Gender}: {percent}%"))

pie(x = Gender$percent, labels = Gender$label, family = "Noto serif")
```

### Marital Status

```{r}
#| fig-width: 11
#| fig-height: 7
Marital <- GAD %>% count(`Marital status`) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{`Marital status`}: {percent}%"))

pie(x = Marital$percent, labels = Marital$label, family = "Noto serif")
```

### Education Level

```{r}
#| fig-width: 11
#| fig-height: 7
Educ <- GAD %>% count(`Education Level`) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{`Education Level`}: {percent}%"))

pie(x = Educ$percent, labels = Educ$label, family = "Noto serif")
```

### Occupation

```{r}
#| fig-width: 11
#| fig-height: 7
Occ <- GAD %>% count(Occupation) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{Occupation}: {percent}%"))

pie(x = Occ$percent, labels = Occ$label, family = "Noto serif")
```

### Residence period

```{r}
#| fig-width: 11
#| fig-height: 7
Res <- GAD %>% count(`Residence period`) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{`Residence period`}: {percent}%"))

pie(x = Res$percent, labels = Res$label, family = "Noto serif")
```

### Nationality

```{r}
#| fig-width: 11
#| fig-height: 7
Nat <- GAD %>% count(Nationality) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{Nationality}: {percent}%"))

pie(x = Nat$percent, labels = Nat$label, family = "Noto serif")
```

### Anxiety

```{r}
#| fig-width: 11
#| fig-height: 7
Anx <- GAD %>% count(GAD_LEVEL) %>% 
  mutate(percent = round(n/sum(n)*100),
         label = glue("{GAD_LEVEL}: {percent}%"))

pie(x = Anx$percent, labels = Anx$label, family = "Noto serif",
    col = brewer.pal(4, "Pastel1"))
```

## piedonut

## Anxiety Vs Gender
```{r}
GAD |> 
  PieDonut(aes(Gender, GAD_LEVEL), family = "Noto serif")
```

## Anxiety Vs Age
```{r}
GAD |> 
  PieDonut(aes(Age, GAD_LEVEL), family = "Noto serif")
```


## Anxiety Vs Marital status
```{r}
GAD |> janitor::clean_names() |> 
  PieDonut(aes(marital_status, gad_level), family = "Noto serif")
```


## Anxiety Vs Education Level
```{r}
GAD |> janitor::clean_names() |> 
  PieDonut(aes(education_level, gad_level), family = "Noto serif")
```


## Anxiety Vs occupation
```{r}
GAD |> janitor::clean_names() |> 
  PieDonut(aes(occupation, gad_level), family = "Noto serif")
```

## Anxiety Vs residence period
```{r}
GAD |> janitor::clean_names() |> 
  PieDonut(aes(residence_period, gad_level), family = "Noto serif")
```