---
title: "Paerasitic Pathogens"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  echo: false
  warning: false
---

```{r}
library(tidyverse)
library(gt)
library(gtsummary)
library(readxl)
Data <- read_excel("REFINED FINNAN DATA.xlsx") |> 
  mutate(`Culture systems` = case_when(`Culture systems` == "Liner pond" ~ "Liner ponds",
                                       `Culture systems` == "Lined earthen pond" ~ "Liner ponds",
                                             `Culture systems` == "Concrete" ~ "Concrete Pond",
                                             .default = `Culture systems`),
         Condition_factor = case_when(`Condition factor` >= 1.6 ~ "Excellent",
                                            between(`Condition factor`,1.4, 1.59) ~ "Good",
                                            between(`Condition factor`,1.2, 1.39) ~ "Fair",
                                            between(`Condition factor`,1.0, 1.19) ~ "Poor",
                                            `Condition factor` <= 0.99 ~ "Extremely Poor"))
  
```

## Weight, lengths and condition factor of sampled fish
```{r}
Data |> select(Weight, `Standard Length`, `Total length`, `Condition factor`) |> 
  tbl_summary( 
              type = everything() ~ "continuous2",
              statistic = all_continuous2() ~ c("{mean}", "{min}", "{max}", "{sd}")) |> 
  modify_header(label = "Variable", stat_0 = "Statistic")
```

## Prevalence of fish parasites as per county
```{r}
Data |> 
  mutate(Total_parasitized = if_else(`Total Parasites` == 0, "No", "Yes")) |> 
  select(Total_parasitized, County) |> 
  tbl_summary(by = County,
              type = Total_parasitized ~ "categorical") |> add_p()
```

## Prevalence of fish parasites as per fish ecosystem
```{r}
Data |> filter(County == "Taita Taveta") |> 
  mutate(Total_parasitized = if_else(`Total Parasites` == 0, "No", "Yes")) |> 
  select(Total_parasitized, Fish) |> 
  tbl_summary(by = Fish,
              type = Total_parasitized ~ "categorical") |> add_p()
```

## Prevalence of fish parasites as per culture system ecosystem
```{r}
Data |> 
  mutate(Total_parasitized = if_else(`Total Parasites` == 0, "No", "Yes")) |> 
  select(Total_parasitized, `Culture systems`) |> 
  tbl_summary(by = `Culture systems`,
              type = Total_parasitized ~ "categorical") |> add_p()
```


## Prevalence of fish parasites as per tilapia species in taveta
```{r}
Data |> filter(County == "Taita Taveta") |> 
  mutate(Total_parasitized = if_else(`Total Parasites` == 0, "No", "Yes")) |> 
  select(Total_parasitized, `Tilapia species`) |> 
  tbl_summary(by = `Tilapia species`,
              type = Total_parasitized ~ "categorical") |> add_p()
```

## ## Prevalence of fish parasites as per condition factor
```{r}
Data |> 
  mutate(Total_parasitized = if_else(`Total Parasites` == 0, "No", "Yes")) |> 
  select(Total_parasitized, Condition_factor) |> 
  tbl_summary(by = Condition_factor,
              type = Total_parasitized ~ "categorical") |> add_p()
```

## Prevalence of different fish parasites
```{r}
Data |> 
  select(`Diplostomum species`:Camallanus) |> 
  mutate_all(~if_else(. == 0, "No", "Yes")) |> 
  tbl_summary() |> modify_header(label = "Parasitic Species", stat_0 = "Prevalence")
```

## Prevalence of different fish parasites as per body condition
```{r}
bind_cols(Data |> select(Condition_factor),
          Data |> 
  select(`Diplostomum species`:Camallanus) |> 
  mutate_all(~if_else(. == 0, "No", "Yes"))) |> 
  tbl_summary(by = Condition_factor,
              percent = "row") |> 
  modify_header(label = "Parasitic Species") |> add_p()
```

## Prevalence of different fish parasites as per County
```{r}
bind_cols(Data |> select(County),
          Data |> 
  select(`Diplostomum species`:Camallanus) |> 
  mutate_all(~if_else(. == 0, "No", "Yes"))) |> 
  tbl_summary(by = County,
              percent = "row") |> 
  modify_header(label = "Parasitic Species") |> add_p()
```


## Prevalence of different fish parasites as per culture system
```{r}
bind_cols(Data |> select(`Culture systems`),
          Data |> 
  select(`Diplostomum species`:Camallanus) |> 
  mutate_all(~if_else(. == 0, "No", "Yes"))) |> 
  tbl_summary(by = `Culture systems`,
              percent = "row") |> 
  modify_header(label = "Parasitic Species") |> add_p()
```

## Prevalence of different fish parasites as per ecosytem
```{r}
bind_cols(Data |> filter(County == "Taita Taveta") |> select(Fish),
          Data |> filter(County == "Taita Taveta") |> 
  select(`Diplostomum species`:Camallanus) |> 
  mutate_all(~if_else(. == 0, "No", "Yes"))) |> 
  tbl_summary(by = Fish,
              percent = "row") |> 
  modify_header(label = "Parasitic Species") |> add_p()
```

## Prevalence of different fish parasites as per tilapia species
```{r}
bind_cols(Data |> filter(County == "Taita Taveta") |> select(`Tilapia species`),
          Data |> filter(County == "Taita Taveta") |> 
  select(`Diplostomum species`:Camallanus) |> 
  mutate_all(~if_else(. == 0, "No", "Yes"))) |> 
  tbl_summary(by = `Tilapia species`,
              percent = "row") |> 
  modify_header(label = "Parasitic Species") |> add_p()
```


# Intensities

## County
```{r}
Data |> 
  select(`Diplostomum species`:Camallanus, County) |> 
  pivot_longer(cols = `Diplostomum species`:Camallanus) |> 
  filter(!value == 0) |> 
  summarise(intensity = round(mean(value), 2),
            sd = round(sd(value), 2),
            .by = c(County, name)) |> 
  unite(Intensity, c(intensity, sd), sep = "±") |> 
  pivot_wider(names_from = County, values_from = Intensity) |> gt()
```

## Body condition
```{r}
Data |> 
  select(`Diplostomum species`:Camallanus, Condition_factor) |> 
  pivot_longer(cols = `Diplostomum species`:Camallanus) |> 
  filter(!value == 0) |> 
  summarise(intensity = round(mean(value), 2),
            sd = round(sd(value), 2),
            .by = c(Condition_factor, name)) |> 
  unite(Intensity, c(intensity, sd), sep = "±") |> 
  pivot_wider(names_from = Condition_factor, values_from = Intensity) |> gt()
```

## Tilpia Species
```{r}
Data |> filter(County == "Taita Taveta") |> 
  select(`Diplostomum species`:Camallanus, `Tilapia species`) |> 
  pivot_longer(cols = `Diplostomum species`:Camallanus) |> 
  filter(!value == 0) |> 
  summarise(intensity = round(mean(value), 2),
            sd = round(sd(value), 2),
            .by = c(`Tilapia species`, name)) |> 
  unite(Intensity, c(intensity, sd), sep = "±") |> 
  pivot_wider(names_from = `Tilapia species`, values_from = Intensity) |> gt()
```

## Culture System
```{r}
Data |> 
  #filter(County == "Taita Taveta") |> 
  select(`Diplostomum species`:Camallanus, `Culture systems`) |> 
  pivot_longer(cols = `Diplostomum species`:Camallanus) |> 
  filter(!value == 0) |> 
  summarise(intensity = round(mean(value), 2),
            sd = round(sd(value), 2),
            .by = c(`Culture systems`, name)) |> 
  unite(Intensity, c(intensity, sd), sep = "±") |> 
  pivot_wider(names_from = `Culture systems`, values_from = Intensity) |> gt()
```

## Ecosystem
```{r}
Data |> 
  filter(County == "Taita Taveta") |> 
  select(`Diplostomum species`:Camallanus, Fish) |> 
  pivot_longer(cols = `Diplostomum species`:Camallanus) |> 
  filter(!value == 0) |> 
  summarise(intensity = round(mean(value), 2),
            sd = round(sd(value), 2),
            .by = c(Fish, name)) |> 
  unite(Intensity, c(intensity, sd), sep = "±") |> 
  pivot_wider(names_from = Fish, values_from = Intensity) |> gt()
```


## County
```{r}
tbl_intensity <- function(group) {
  Data |> select(`Diplostomum species`:Camallanus, {{group}}) |> 
  pivot_longer(cols = `Diplostomum species`:Camallanus) |> 
  filter(value > 0) |> 
  summarise(intensity = mean(value),
            .by = c(name, {{group}})) |> 
  pivot_wider(names_from = name, values_from = intensity) |> 
  tbl_summary(by = {{group}},
              type = everything() ~ "continuous",
              statistic = all_continuous() ~ "{mean}",
              missing = "no") |> 
  add_p()
}

tbl_intensity(County)

```
## Body Cndition
```{r}
tbl_intensity(Condition_factor)
```

## Culture systems
```{r}
tbl_intensity("Culture systems")
```

## Tilapia species
```{r}
Data <- Data |> filter(County == "Taita Taveta")
tbl_intensity("Tilapia species")
```

## SPECIES ECOSYSTEM
```{r}
tbl_intensity(Fish)
```

