---
title: "Publication"
author: "Mueke"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  echo: false
  warning: false
---

```{r}
library(tidyverse)
library(readxl)
library(gt)
library(gtsummary)
farmers <- read_excel("farmers.xlsx") |> 
   mutate(Age = case_when(Age == "option_2" ~ "36-50",
                         Age == "36_50" ~ "36-50",
                         Age == "_60" ~ "Above 60",
                         Age == "18_35" ~ "18-35",
                         Age == "51_60" ~ "51-60"),
         Education_level = fct_relevel(Education_level, c("no_schooling", "primary_education",
                                                          "secondary_education",
                                                          "college_training__certificate__diploma",
                                                          "bachelor_degree_and_above")),
         Marital_status = fct_relevel(Marital_status, c("single","married", "windowed", "separated",
                                                        "divorced")),
         Household_income = case_when(Household_income == "_5000" ~ "Below 5000",
                                      Household_income == "_20000" ~ "10000-20000",
                                      Household_income == "10000_20000" ~ "10000-20000",
                                      Household_income == "5000_10000" ~ "5000-10000"))

biologicals <- read_excel("farmers.xlsx", 
    sheet = "Sheet2")
```

## Social Demographics

```{r}
demographics <- farmers |> 
  select(Age, Education_level, Marital_status, Household_income, Head_of_household,
         farming_years) 
demographics |> 
  tbl_summary(statistic = farming_years ~ "{mean}",
              missing = "no") |> 
  modify_header(label = '**Social Demographics**',
                stat_0 = 'Frequency') |> 
  bold_labels()
```

## Biologicals Knowledge

```{r}
Knowledge <- bind_cols(farmers |> select(pesticide_purpose, fertilizer_purpose, knows_pesticeide_type, knows_pesticide_content, human_healthy_effects,
                  animal_healthy_effects, environment_effects, crop_production_effects, mitigation) |> 
  mutate(pesticide_purpose = case_when(str_detect(pesticide_purpose, "to_reduce_crop_loss|to_fulfil_market_demand") ~ "Sufficient",
                                       .default = "Insufficient"),
         fertilizer_purpose = case_when(str_detect(fertilizer_purpose, "to_improve_production") ~ "Sufficient",
                                       .default = "Insufficient"),
         knows_pesticeide_type = case_when(str_detect(knows_pesticeide_type, "yes") ~ "Sufficient",
                                       .default = "Insufficient"),
         knows_pesticide_content = case_when(str_detect(knows_pesticide_content, "yes") ~ "Sufficient",
                                       .default = "Insufficient"),
         human_healthy_effects = if_else(human_healthy_effects == "yes", "Sufficient", "Insufficient"),
         animal_healthy_effects = if_else(animal_healthy_effects == "yes", "Sufficient", "Insufficient"),
         environment_effects = if_else(environment_effects == "yes", "Sufficient", "Insufficient"),
         crop_production_effects = if_else(crop_production_effects == "yes", "Sufficient", "Insufficient"),
         human_healthy_effects = if_else(human_healthy_effects == "yes", "Sufficient", "Insufficient"),
         mitigation = if_else(str_detect(mitigation, "i_do_nothing"), "Insufficient", "Sufficient")),
  biologicals |> select(input_choice) |> 
        mutate(input_choice = if_else(str_detect(input_choice, "biologicals"), "Sufficient", "Insufficient")))

Knowledge |> tbl_summary(missing = "no") |> 
  modify_header(label = "Knowledge Statement", stat_0 = "Knowledge Prevalence")
```

### Knowledge scores
```{r}
Know_scores <- Knowledge |> 
  mutate_all(~if_else(. == "Sufficient", 1, 0)) |> 
  rowwise() |> 
  mutate(Knowledge_score = sum(c_across(1:10), na.rm = T),
         Knowledge_level = if_else(Knowledge_score >= 5, "Sufficient", "Insufficient")) |> 
  select(11,12)
  
```

## Biologicals Attitude

```{r}
Attitude <- biologicals |> select(biologicals_advantageous:recommend_biologicals) |> 
  mutate_at(.vars = c("Biologicals_increase_yield_of_crops", "Biologicals_increase_incomes",
                      "biologicals_advantageous", "Biologicals_can_produce_healthy_food",
                      "recommend_biologicals", "government_support_biologicals"),
            ~ if_else(str_detect(., "agree|strongly_agree"),"Preferrable", "Unpreferrable")) |> 
  mutate_at(.vars = c("Biologicals_risk_free", "biologicals_environmentally_safe",
                      "biologicals_use_unconditional"),
            ~if_else(str_detect(., "strongly_disagree|disagree"), "Preferrable", "Unpreferrable")) |> 
  select("Biologicals_increase_yield_of_crops", "Biologicals_increase_incomes",
         "biologicals_advantageous", "Biologicals_can_produce_healthy_food",
         "recommend_biologicals", "government_support_biologicals","Biologicals_risk_free",
         "biologicals_environmentally_safe",
         "biologicals_use_unconditional")

Attitude |> tbl_summary(missing = "no") |> 
  modify_header(label = "Attitude Statement", stat_0 = "Attitude Prevalence")
  
```

## Attitude scores
```{r}
Att_scores <- Attitude |> mutate_all(~if_else(. == "Preferrable", 1, 0)) |> 
  rowwise() |> 
  mutate(Attitude_score = sum(c_across(1:9), na.rm = T),
         Attitude_level = if_else(Attitude_score >= 4.5, "Preferrable", "Unpreferrable")) |> 
  select(10,11)
```

## Biologivals Practices

```{r}
Practices <- bind_cols(farmers |> select(usse_agricultural_inputs, agricultural_input, bought_safer_pesticide) |> 
  mutate(usse_agricultural_inputs = if_else(usse_agricultural_inputs == "yes", "Valid", "Invalid"),
         agricultural_input = if_else(str_detect(agricultural_input, "none|caco3|compost"), "Invalid", "Valid"),
         bought_safer_pesticide = if_else(bought_safer_pesticide == "yes", "Valid", "Invalid")),
  biologicals |> select(specific_names, use_biologicals,used_biologicals_last_farming_season,biologicals_use_frequency) |> 
  mutate(specific_names = if_else(str_detect(specific_names, "not_applicable"), "Invalid", "Valid"),
         used_biologicals_last_farming_season = if_else(used_biologicals_last_farming_season == "yes", "Valid", "Invalid"),
         use_biologicals = if_else(use_biologicals == "yes", "Valid", "Invalid"),
         biologicals_use_frequency = if_else(str_detect(biologicals_use_frequency, "every_season|some_seasons"), "Valid", "Invalid")) |> 
  rename(Biological_use_times = specific_names))

Practices |> tbl_summary(missing = "no") |> 
  modify_header(label = "Practices Statement", stat_0 = "Practices Prevalence")
```

## Practices scores
```{r}
pract_scores <- Practices |> mutate_all(~if_else(. == "Valid", 1, 0)) |> 
  rowwise() |> 
  mutate(Practices_score = sum(c_across(1:7), na.rm = T),
         Practices_level = if_else(Practices_score >= 3.5, "Invalid", "Valid")) |> 
  select(8,9)
```

## KAP
```{r}
KAP <- bind_cols(demographics,
          Know_scores,
          Att_scores,
          pract_scores)

KAP
```

## Know
```{r}
KAP |> select(Knowledge_level) |> 
  tbl_summary()
```

## Attitude
```{r}
KAP |> select(Attitude_level) |> 
  tbl_summary()
```

## Practices
```{r}
KAP |> select(Practices_level) |> 
  tbl_summary()
```

## Association between Knowledge and social demographics
```{r}
KAP |> 
  select(1:6,8) |> 
  tbl_summary(by = Knowledge_level, missing = "no", 
              statistic = farming_years ~ "{mean} ({sd})") |> 
  add_p(test = all_categorical() ~ "chisq.test")
```

## Association between practices and social demographics
```{r}
KAP |> 
  select(1:6,Practices_level) |> 
  tbl_summary(by = Practices_level, missing = "no", 
              statistic = farming_years ~ "{mean} ({sd})") |> 
  add_p(test = all_categorical() ~ "chisq.test")
```


## Association between attitude and social demographics
```{r}
KAP |> 
  select(1:6,Attitude_level) |> 
  tbl_summary(by = Attitude_level, missing = "no", 
              statistic = farming_years ~ "{mean} ({sd})") |> 
  add_p(test = all_categorical() ~ "chisq.test")
```


## Knowledge Regression
```{r}
Model_data <- KAP |> 
  mutate(Knowledge_level = if_else(Knowledge_level== "Sufficient" , 1,0),
         Attitude_level = if_else(Attitude_level == "Preferrable", 1, 0),
         Practices_level = if_else(Practices_level == "Valid", 1, 0))
model <- glm(Knowledge_level ~ Age + Education_level + Marital_status + Household_income + farming_years, data = Model_data,
             family = binomial(link = "logit"))

model |> tbl_regression()

```

## Attitude Regression
```{r}
model_1 <- glm(Attitude_level ~ Age + Education_level + Marital_status + Household_income + farming_years, data = Model_data,
             family = binomial(link = "logit"))

model_1 |> tbl_regression()
```

## Practices Regression
```{r}
model_2 <- glm(Practices_level ~ Age + Education_level + Marital_status + Household_income + farming_years, data = Model_data,
             family = binomial(link = "logit"))

model_2 |> tbl_regression()
```

## Correlation
```{r}
corrdata <- KAP |> select(Knowledge_score, Attitude_score, Practices_score)

library(GGally)
library(corrr)

ggcorr(corrdata, label = T)

corrdata |> correlate() |> gt()
```


